---
title: "Brief - Mark Fletcher"
author: "Patrick Leslie"
date: "09/12/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F)
library(tidyverse)
library(readr)
library(kableExtra)
acts <- read.csv("outputs/a_file_2_acts_plus_event.csv")
parl_dates <- read.csv("inputs/parl_dates.csv")


parliament_sum<-acts%>%group_by(parl_enacted)%>%
  summarise(amended_same_term=mean(amended_same_term),
            repealed_same_term=mean(repealed_same_term),
            altered_same_term=mean(altered_same_term))%>%
  left_join(parl_dates%>%select(election_date,name), by=c('parl_enacted'='name'))
parliament_sum$election_date<-as.Date(parliament_sum$election_date, format = "%d.%m.%Y")

parliament_sum<-parliament_sum%>%
  arrange(election_date)

parliament_sum<-parliament_sum[,c(1,5,2:4)]

parliament_sum[,3:5]<-parliament_sum[,3:5]*100


```

## Role Background -- Are early amendments evidence of poor quality legislation?

Quality of legislation is (we know) an extremely tricky thing to measure but we've done some early work scraping the Federal Register of Legislation picking up all Acts and amendments since 1901, looking for ways to infer quality from easily available quant data. One candidate such measure is to look at Acts of Parliament that were amended or repealed before the end of a parliamentary term, implying perhaps that the government realises it's made a mistake and needs to change something.

We have some ideas about what might be causing the big peaks and troughs and some of this historical reasoning is quite interesting but the next step is to look more granularly at the legislation to see what patterns emerge from the bottom up. Reasons for quick amendments might be institutional (senate control), personnel (who is PM minister or drafter), more effective scrutiny, political or economic crises, or simply administrative changes caused by omnibus amending acts.

```{r, fig.cap="Percentage of Acts amended or repealed before the end of the parlimentary term in which the act was amended, 1901-2021."}


ggplot(parliament_sum, aes(election_date, altered_same_term))+
  geom_step() + ylab("Amended or Repealed \nsame term as enacted (%)") + xlab('')+
  scale_y_continuous(limits = c(0,50)) + theme_minimal() + 
  scale_x_date(date_breaks="5 year", date_labels = "%Y")+
  theme(axis.text.x = element_text(angle = 90))+
  theme(axis.text.x = element_text(vjust = 0.5))
```

## The task

The task is to determine how the following acts came to be amended or repealed before the end of the term in which they were enacted. The secondary (though ultimate) goal of the exercise is to determine why this phenomenon is on the increase, and whether or not it has anything to do with the quality of legislation.

50 acts, randomly sampled from five periods. Write a simple one-pager sumarising each period and what happened with the respective acts. Give an idea of how each act changed in tabular or bullet point form. If there are any general thoughts about what you can infer from these acts about the period, that would be excellent too.

```{r}

set.seed(123)

early<-sample(acts$id_principal[acts$date_enacted<as.Date("1961-12-09")&acts$altered_same_term==1], 10, replace = F)


mid_1<-sample(acts$id_principal[acts$date_enacted>as.Date("1961-12-09")& acts$date_enacted<as.Date("1983-03-05")&acts$altered_same_term==1], 10, replace = F)


lab_83_96 <-sample(acts$id_principal[acts$date_enacted<as.Date("1996-03-02")& acts$date_enacted>as.Date("1983-03-05")&acts$altered_same_term==1], 10, replace = F)

howard <-sample(acts$id_principal[acts$date_enacted<as.Date("2007-11-24")& acts$date_enacted>as.Date("1996-03-02")&acts$altered_same_term==1], 10, replace = F)

recent <-sample(acts$id_principal[acts$date_enacted<as.Date("2019-05-18")& acts$date_enacted>as.Date("2007-11-24")&acts$altered_same_term==1], 10, replace = F)

tab<-bind_cols(`Early 1901-1961` = early,
          `60s and 70s` = mid_1, 
          `Labor 1983-96` = lab_83_96, 
          `Coalition 1996-2007`= howard, 
          `Mixed 2007-2019`=recent)

tab%>%kbl(caption = "Commonwealth Law ID numbers of Acts to Invesigate", booktabs=T,
  escape=FALSE)%>%
  kable_styling(latex_options = "HOLD_position")

```

## Details

ANU Research Officer 5, reporting to Professor Keith Dowding (supervisor) and Dr Patrick Leslie. Initial contract for **up to 20** hours RA time.
